---
title: "Class 11: Structural Bioinformatics 1"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## The PDB database for biomolecular structure data
>Q1: determine the percentage of structures solved by X-Ray and Electron Microscopy.
      X-ray is 89%. Electron Microscopy is 2%
>Determine what proportion of structures are protein?
      92.71%

```{r}
data<-read.csv("Data Export Summary.csv",row.names=1)
data
```

Total number of entries
```{r}
sum(data$Total)
```
Proportion of entries from each ethod
```{r}
round((data$Total/sum(data$Total))*100)
```

Proportion that are protein
```{r}
round((sum(data$Proteins)/sum(data$Total))*100,2)
```

##HIV-Pr Structure

##Here we will read the 1HSG PDB structure and select the protein component and write out a new protein-only PDB format file. We then do the same for the ligand (i.e. known drug molecule) creating a ligand-only PDB file.
```{r}
library(bio3d)
pdb<-read.pdb("1hsg.pdb")
```
#MK1 is the compound
```{r}
ligand<-atom.select(pdb,"ligand",value=TRUE)
write.pdb(ligand,file="1hsg_ligand.pdb")
```



