---
title: "class09"
author: "Yu Yan Kwan"
date: "10/29/2019"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

##Data input
# Complete the following code to input the data and store as wisc.df
```{r}
wisc.df<-read.csv("WisconsinCancer.csv")
```

Not that the 'id' and 'diagnosis' clumns will not be used for most of the following steps
We have 'r nrow(wisc.df)' samples in this dataset

```{r}
nrow(wisc.df)
```

How many benign(not cancerous) and malignant (cancerous) samples do we have in the dataset
```{r}
table(wisc.df$diagnosis)
```

# We want to delete the 'id' and'diagnostics' columns 
# Convert the features of the data: wisc.data
```{r}
wisc.data<-as.matrix(wisc.df[,3:32])
head(wisc.data)
```

# Set the row names of wisc.data
```{r}
row.names(wisc.data) <- wisc.df$id
head(wisc.data)
```

#Store the diagnosis for reference in the future as a seperate vector
#Create diagnosis vector by completing the missing code 
```{r}
diagnosis<-wisc.df$diagnosis
```

Questions:
1. How many observations are in this dataset? 
```{r}
nrow(wisc.df)
```

2. How many variables/features in the data are suffixed with _mean?
```{r}
table(wisc.df$diagnosis)
```
```{r}
dianosis<-as.numeric(wisc.df$diagnosis=="M")
```

3.  How many of the observations have a malignant diagnosis?
```{r}
grep("_mean",colnames(wisc.df),value=TRUE)
```

#length() helps count how matches we have.
```{r}
length(grep("_mean",colnames(wisc.df)))
```



#Section 2
#Perform PCA on wisc.data
#It is important to check if the data need to be scaled before performing PCA. Recall two common reasons for scaling data include:
  #The input variables use different units of measurement.
  #The input variables have significantly different variances.
```{r}
round(colMeans(wisc.data),3)
```

```{r}
apply(wisc.data,2,sd)
```

These values look very different so I will use "scale=FALSE"

# Perform PCA on wisc.data by completing the following code
```{r}
wisc.pr <- prcomp((wisc.data),scale=TRUE)
```

#look at the summary of results
```{r}
summary(wisc.pr)
```


#Color by cancer/non-cancer
```{r}
plot(wisc.pr$x[,1],wisc.pr$x[,2],col=diagnosis)
```

Questions:
4. From your results, what proportion of the original variance is captured by the first principal components (PC1)?

```{r}
x<-summary(wisc.pr)
x$importance[,"PC1"]
```


```{r}
biplot(wisc.pr)
```



Let's see if PCA improves or degrades the performanc of hierarchical clustering
```{r}
wisc.pr.hclust<-hclust(dist(wisc.pr$x[,1:7]))
```

Using the minimum number of principal components required to describe at least 90% of the variability in the data, create a hierarchical clustering model with the linkage method="ward.D2". We use Ward’s criterion here because it is based on multidimensional variance like principal components analysis. Assign the results to wisc.pr.hclust.
```{r}
plot(wisc.pr.hclust)
```


```{r}
grps<-cutree(wisc.pr.hclust, k=2)
table(grps)
```
```{r}
table(grps,diagnosis)
```

```{r}
plot(wisc.pr$x[,1:2], col=diagnosis)
```

##Section 7 Prediction
#We will use the predict() function that will take our PCA model from before and new cancer cell data and project that data onto our PCA space.
```{r}
new<-read.csv("new_samples.csv")
npc<-predict(wisc.pr,newdata=new)
npc
```
#plot these new samples on our PC1 vs PC2 plot
```{r}
plot(wisc.pr$x[,1:2], col=diagnosis)
points(npc[,1], npc[,2], col="blue", pch=16, cex=3)
text(npc[,1], npc[,2], c(1,2), col="white")
```

**Q17** Which of these new patients should we prioritize for follow up based on your results?

