---
title: "Class 08 Machine Learning"
author: "Yu Yan Kwan"
date: "10/24/2019"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

##K means clustering
```{r}
#Generate some example data for clustering
tmp<-c(rnorm(30,-3),rnorm(30,3))
x<-cbind(x=tmp,y=rev(tmp))
plot(x)
```


```{r}
k<-kmeans(x,centers=2,nstart=20)
```

```{r}
k
```

Q. How many points are in each cluster?
Q. What "component" of your result object details
  -cluster size?
  -cluster assignment/membership?
  -cluster center?
  
```{r}
k$size
```

```{r}
k$cluster
```

```{r}
k$centers
```

Plot x colored by tehe kmeans cluster assignment and add cluster centers as blue points
#pch makes the point a square
```{r}
plot(x,col=k$cluster)
points(k$centers,col="blue",pch=15)
```



##Hierarchical clutering in R
The "hclust()" function requires a distance matrix as input. You can get this from the "dist()" function. Often see "hclust()" and "dist()" together. 

```{r}
# First we need to calculate point (dis)similarity
#   as the Euclidean distance between observations
dist_matrix <- dist(x)
# The hclust() function returns a hierarchical
#  clustering model
hc <- hclust(d = dist_matrix)
# the print method is not so useful here
hc

```

```{r}
plot(hc)
abline(h=6,col="red")
grps<-cutree(hc,h=6)
```


#Step1. Generate some example data for clustering
```{r}
x <- rbind(
  matrix(rnorm(100, mean=0, sd = 0.3), ncol = 2),   # c1
  matrix(rnorm(100, mean = 1, sd = 0.3), ncol = 2), # c2
  matrix(c(rnorm(50, mean = 1, sd = 0.3),           # c3
           rnorm(50, mean = 0, sd = 0.3)), ncol = 2))
colnames(x) <- c("x", "y")
```

#Step2. Plot the data without clustering
```{r}
plot(x)
```

# Step 3. Generate colors for known clusters
#         (just so we can compare to hclust results)
```{r}
col <- as.factor( rep(c("c1","c2","c3"), each=50) )
plot(x, col=col)
```
#There are some black, green and red point overlapping. 


Q. Use the dist(), hclust(), plot() and cutree() functions to return 2 and 3 clusters
Q. How does this compare to your known 'col' groups?

```{r}
#clustering
hc<-hclust(dist(x))
#draw tree
plot(hc)
abline(h=2,col="red")
#cut the tree into clusters/groups
groups<-cutree(hc,h=3)
groups
```


Plot the data colored by the hclust result with k=3
```{r}
plot(x,col=groups)
```

How many points in each cluster?
```{r}
table(groups,col)
```
#the table tells you that cluster 1 has 94 members, cluster 2 has 56 members 


```{r}
mydata <- read.csv("https://tinyurl.com/expression-CSV",
row.names=1)

head(mydata)
```

PCA with the **prcomp()* function
```{r}
## lets do PCA
pca <- prcomp(t(mydata), scale=TRUE)
## See what is returned by the prcomp() function  attributes(pca)
attributes(pca)
## A basic PC1 vs PC2 2-D plot  plot(pca$x[,1], pca$x[,2])
plot(pca$x[,1], pca$x[,2])
## Precent variance is often more informative to look at  
pca.var <- pca$sdev^2
pca.var.per <- round(pca.var/sum(pca.var)*100, 1)
pca.var.per

```

```{r}
pca.var <- pca$sdev^2
pca.var.per <- round(pca.var/sum(pca.var)*100, 1)

barplot(pca.var.per, main="Scree Plot",
xlab="Principal Component", ylab="Percent Variation")

```


```{r}
plot(pca$x[,1], pca$x[,2], 
     col=c("red","red","red","red","red","blue","blue","blue","blue","blue"))
```

##Hands on 
```{r}
x <- read.csv("UK_foods.csv",row.names=1)
x
```

# Note how the minus indexing works
```{r}
rownames(x) <- x[,1]
x <- x[,-1]
head(x)
```

```{r}
dim(x)
```

```{r}
barplot(as.matrix(x), beside=T, col=rainbow(nrow(x)))
pairs(x, col=rainbow(10), pch=16)
```

```{r}
pca <- prcomp( t(x) )
summary(pca)
```

