---
title: "Class 14"
author: "Yu Yan Kwan"
date: "11/12/2019"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

##Inpur dataset
Input countData and colData (metadata)
```{r}
counts <- read.csv("airway_scaledcounts.csv", stringsAsFactors = FALSE)
metadata <-  read.csv("airway_metadata.csv", stringsAsFactors = FALSE)
```

```{r}
head(counts)
```

```{r}
head(metadata)
```

#How many genes do we have in this dataset?
```{r}
nrow(counts)
```

#How many experiments do we have?
```{r}
ncol(counts)
```

```{r}
nrow(metadata)
```

#We want to know if there is a difference in expression values for control (non-drug) vs treated (drug added cell lines).
#First step is to find which experiments were the control experiments and then get the average values accross all control experiments. Then we will do the same for drug treated. 

```{r}
control <- metadata[metadata[,"dex"]=="control",]
control$id
```

#Now calculate the mean values accorss these control columns of **countdata**
```{r}
control.mean<-rowSums(counts[,control$id])/length(control$id)
names(control.mean)<-counts$ensgene
```

#Now we find the treated experiments.
```{r}
treated<-metadata[metadata[,"dex"]=="treated",]
treated$id
```

#Now calculate the mean values accross the treated columns of **countdata**
```{r}
treated.mean<-rowSums(counts[,treated$id])/length(treated$id)
names(treated.mean)<-counts$ensgene
```

#We will combine our meancount data for bookkeeping purposes.
```{r}
meancounts<-data.frame(control.mean,treated.mean)
```

#Plot control vs treated
```{r}
plot(meancounts$control.mean,meancounts$treated.mean)
```

#This will be better on a log scale
```{r}
plot(meancounts$control.mean,meancounts$treated.mean,log="xy")
```


#Now we want to calculate log2foldchange
```{r}
meancounts$log2fc <- log2(meancounts[,"treated.mean"]/meancounts[,"control.mean"])
head(meancounts)
```

#We want to delete data which have 0 in it. 
```{r}
zero.vals <- which(meancounts[,1:2]==0, arr.ind=TRUE)

to.rm <- unique(zero.vals[,1])
mycounts <- meancounts[-to.rm,]
head(mycounts)
```

#A common threshold used for calling something differentially expressed is a log2(FoldChange) of greater than 2 or less than -2. Let’s filter the dataset both ways to see how many genes are up or down-regulated.
```{r}
up.ind <- mycounts$log2fc > 2
down.ind <- mycounts$log2fc < (-2)
```

```{r}
sum(up.ind)
```

```{r}
head(mycounts[up.ind,])
```



##DESeq2 Analysis
#DESeq2 is an R package for analyzing count-based NGS data like RNA-seq. It is available from Bioconductor. Bioconductor is a project to provide tools for analyzing high-throughput genomic data including RNA-seq, ChIP-seq and arrays.
```{r}
library(DESeq2)
```

```{r}
dds <- DESeqDataSetFromMatrix(countData=counts, 
                              colData=metadata, 
                              design=~dex, 
                              tidy=TRUE)
dds
```

#Run DESeq2
```{r}
dds<-DESeq(dds)
```

```{r}
res <- results(dds)
res
```


#Summary plot: aka Volcano plots
#This figure will combine both Fold Change and the p-value into one overview figure indicating the porportion of genes with large scale significant differences in their expression.
```{r}
#Dull all-black plot
plot(res$log2FoldChange,-log(res$padj))
abline(v=c(-2,2),col="gray",lty=2)
abline(h=-log(0.05),col="gray",lty=2)
```

```{r}
#Add some color please
mycols<-rep("gray",nrow(res))

#make points wit +2 -2 fold change blue
mycols[abs(res$log2FoldChange)>2]<-"blue"

#make points below p-value cutoff gray
mycols[(res$padj>0.05)]<-"gray"
plot(res$log2FoldChange,-log(res$padj),col=mycols)
abline(v=c(-2,2),col="gray",lty=2)
abline(h=-log(0.05),col="gray",lty=2)
```

#Save our results for next day
```{r}
write.csv(res,file="expression_results.csv")
```

