---
title: "Class 7 R function and packages"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```



## R functions revisited
```{r}
source("http://tinyurl.com/rescale-R")
```

```{r}
rescale(1:10)
```

```{r}
rescale(c(1,10,5,NA,6))
```

is.numeric(1:10)
!is.numeric(c(1,5,NA,10,"barry"))


```{r}
rescale2(c(1:10),"barry")
```


## A new function called both_na()
Write a function to find where there are NA elements in two input vectors. 
First make some simple input where I know the answer
```{r}
x<-c(1,2,NA,3,NA)
y<-c(NA,3,NA,3,4)
```

Looked online and found the **is.na** function
```{r}
is.na(x)
```

and the **which()** function tells me where the TRUE values are
```{r}
which(is.na(x))
```


```{r}
which(is.na(y))
```
The AND function requires two input TRUE to give a TRUE output
```{r}
is.na(x)&is.na(y)
```

Taking the **sum()** of TRUE FALSE vector will tell me how many TRUE elements I have.This is my working snippet!
```{r}
sum(is.na(x)&is.na(y))
```


```{r}
sum(c(TRUE,TRUE,FALSE,TRUE))
```


Now turn it into a function
```{r}
both_na<- function(x,y){
sum(is.na(x)&is.na(y))
}
```

```{r}
both_na(x,y)
```

Test function 
```{r}
x<-c(NA,NA,NA)
y1<-c(1,NA,NA)
y2<-c(1,NA,NA,NA)
```

```{r}
both_na(x,y2)
```

```{r}
x2<-c(NA,NA)
both_na(x2,y2)
```


```{r}
x<-c(NA,NA,NA)
y3<-c(1,NA,NA,NA,NA,NA,NA)
```


```{r}
both_na2<- function(x,y){
sum(is.na(x)&is.na(y))
}
```


```{r}
both_na2<-function(x,y){
  if(length(x)!=length(y)){
    stop("inputs x and y should be the same length!")
  }
  sum(is.na(x)&is.na(y))
}
```

```{r}
both_na2(x,y3)
```



Practice
```{r}
#student1
s1<-c(100,100,100,100,100,100,100,90)
#student2
s2<-c(100,NA,90,90,90,90,97,80)

mean(s1[-which.min(s1)])
```

```{r}
which.min(s2)
s2[-which.min(s2)]
mean(s2[-which.min(s2)],na.rm=TRUE)
```

We have our working code now turn it into a first function
```{r}
grade<-function(x){
  if(any(is.na(x))){
    warning("Student is missing a homework")
  }
  mean(x[-which.min(x)],na.rm=TRUE)
}
```

```{r}
grade(s2)
```

```{r}
grade2<-function(x){
  if(any(is.na(x))){
    warning("Student is missing a homework")
  }
  sum(x[-which.min(x)],na.rm=TRUE)/(length(x)-1)
}
```


```{r}
grade(s2)
```
```{r}
url<-"http://tinyurl.com/gradeinput"
hw<-read.csv(url,row.names=1)
```

```{r}
apply(hw,1,grade)
```


```{r}
install.packages("blogdown")
```

